Test-driven Knowledge Graph Construction

### 2023-05-08 doc/paper4.tex
More editing

%%%Turtle
\lstdefinelanguage{turtle}
{
  identifierstyle=\color{afterPrefix},    %afterPrefix
  morecomment=[n][\color{iri}]{<}{>}, % iris
  morestring=[b][\color{black}]{\"},  % strings
  % -------------------------- prefixes
  classoffset=1,
  keywordstyle=\color{prefix},
  morekeywords={acte, dcterms, owl, rdfs, rdf, schema, reg, skos, xsd, sh},
  % -------------------------- rdfType
  classoffset=2,
  keywordstyle=\color{purple},
  morekeywords={a},
  % -------------------------- language
  classoffset=3,
  keywordstyle=\color{orange},
  morekeywords={@en}
}

@@

    \item Wrap the queries as SPARQL-based SHACL constraints to automate their execution. SHACL defines a target\footnote{\urlshacltargets} of each constraint. In our context, a target sets the scope of a CQ. The scope of the data graph validated by a given constraint may either encompass the entire knowledge graph or cover its subset. For instance, existentially quantified queries checking the complete data graph can target it by using \texttt{sh:targetNode []}. It is also possible to rewrite a part of a SPARQL query translating a CQ as target selection in SHACL. Alternatively, the prerequisite part of the query can be represented as a SPARQL-based target \cite{SHACLAF2017}. An~entity-centric partitioning can be implemented by extracting concise bounded descriptions~\cite{Stickler2005} of the targeted entities. Some data sources for the knowledge graph may natively partition data to independent subsets, such as API responses or messages from a queue, that can be effectively validated by specific subsets of SHACL constraints. SHACL also defines the criteria of correctness of the SPARQL queries. An \texttt{sh:ask} query is expected to return the boolean \texttt{true}, while any results produced by an sh:select query are treated as violations. Specific expected results can be either hard-coded into the queries or specified via \texttt{sh:hasValue} in case a single value is expected. When including the expected results in SPARQL-based constraints, we may run into the limitations of SHACL due to pre-binding of variables.\footnote{\urlshaclprebinding} For example, the \texttt{VALUES} clause, which could represent results of \texttt{SELECT} queries, is not allowed.

@@

    \item Validate the knowledge graph under development with the SHACL constraints. If the validation fails, resolve the reported violations by fixing the artifacts created in the previous steps. Fix the most primary artifact causing the violations. For instance, data transformation shall not work around an insufficiently expressive ontology. Similarly, imposing a more expressive data model on the knowledge graph might reveal errors in its source data. In this way, the finer structure of its ontology makes the previously hidden data quality issues visible. Some feedback may even indicate ill-formed CQs in need of revision.
    \item Refactor the artifacts. The SHACL specification of the knowledge graph is a key artifact to refactor. It can serve as a formal specification of the anticipated use of the ontology. Moreover, SHACL can transcend a single ontology and specify how to combine terms reused from multiple ontologies. SHACL core constraints can be induced from SPARQL-based constraints to represent the frequently queried patterns in data. For instance, you can infer universally quantified axioms represented as SHACL constraints from a sample of existentially quantified axioms represented as SPARQL queries. In particular, SHACL can represent the expected relations in data. For example, more complex or distant relations might require the expressivity of SHACL to be represented, using property paths or SPARQL graph patterns. Even though SHACL does not support conditional constraints directly, they can be rewritten according to the inference rule of material implication ${P\to Q\Leftrightarrow \neg P\lor Q}$.

@@

\section{Non-functional Requirements}
Note that the above-described method tests if a knowledge graph meets the given user requirements. It does not evaluate how well these requirements are satisfied. Absence of errors does not imply that the knowledge graph is sound. Therefore, the functional user requirements should be complemented by non-functional requirements describing the desired qualities of the solution. These requirements can be checked and refactored by using alternative sources of feedback, such as:

\begin{description}
    \item[Code review] \hfill \\
    Elicit expert insight from ontology and data engineers.
    \item[User feedback] \hfill \\
    The results produced by SPARQL queries formalizing the CQs can be judged to be incorrect by users. User feedback can identify incorrect interpretation of CQs or incorrect assumptions. It might prompt revisiting any of the previously described steps or lead to adding more CQs.
    \item[Usability] \hfill \\
    Usability may manifest in developer experience of writing SPARQL queries on a knowledge graph. In particular, complexity and verbosity of SPARQL queries may be considered. In this way, usability of the queries indirectly reflects the usability of the ontology. For example, verbosity may be caused by duplicate data that can be abstracted to the ontology, while complexity can be reduced by introducing ontological shortcuts. Verbose SPARQL queries may also suggest a need for introducing abstraction or additional axioms into the ontology. In fact, some CQs can be answered directly using the knowledge formalized in the ontology, since SHACL expects the validated data graph\footnote{\urlschaclgraphdata} to include the ontologies it populates.
    \item[Performance] \hfill \\
    Performance can be evaluated indirectly using the execution time of SHACL validation. This time is proportional to the execution times of the SPARQL queries the constraints include.
\end{description}

The non-functional requirements can also be covered by tests, such as query performance tests. Such tests can provide additional feedback informing the test-driven development.

Using the above-mentioned feedback is a balancing act of multi-objective optimization. For example, performance feedback may be incompatible with the feedback about verbosity. CQs may pose conflicting requirements. Addressing the feedback thus requires making informed trade-offs.

\section{Limitations}
We are aware of several limitations of the proposed method. Here we recount some of them and suggest how to remedy them. The method may cause the developed artifacts outlined in the Figure \ref{fig:process} to overfit the user requirements in scope, which would hinder the reuse of the artifacts. This shortcoming can be remedied by including more CQs or focusing on reusability during refactoring. User requirements for the knowledge graph under construction can be more complex than what SPARQL can express. One way around it is formalizing the requirements in a more expressive programming language that extends SHACL, such as in \cite{SHACLJS2017}. Ultimately, in order to ameliorate the limitations of this method, it is best combined with other methods, such as those for ontology design (e.g., Kendall and McGuinness \cite{Kendall2019}).

@@

The data was collected manually from several regulatory bodies, such as SÚKL\footnote{\url{https://www.sukl.cz/prehled-testu-k-diagnostice-onemocneni-covid-19}} and EU HSC\footnote{\url{https://health.ec.europa.eu/health-security-and-infectious-diseases/crisis-management/covid-19-diagnostic-tests_en}}, and was combined with performance evaluations of the tests, such as their sensitivity and specificity, that came from independent studies from various countries.
The data was collected into an XML file.
The final destination of the data was originally twofold.
One was the actual portal that provided a graphical visualization of various properties of the tests (\url{https://covidtesty.vse.cz/english/test-evaluation-older-data/}).
The other was a dashboard allowing to perform analyses online.
While the XML data was already structured, it lacked any ontological grounding, and some of its features were tuned towards presentation aspects within the website.

@@

\textbf{Step 1:} The first step of the proposed method is gathering user requirements. We started with capturing user requirements formulated as CQs, such as \textit{"What is the sensitivity of given tests according to their manufacturers?"} We came up with 19 CQs in total, covering basic information look-up and more complex analytical questions. The complete list of the CQs is available online.\footnote{\url{https://github.com/KIZI/antigen-covid-tests-knowledge-graph/wiki/Competency-questions}}

@@

\textbf{Step 4:} Having a minimum viable ontology, we were able to translate the CQs into SPARQL queries. Each question was translated manually.

\textbf{Step 5:} We wrapped the SPARQL queries as SHACL constraints, such as in the example Listing \ref{lst:example-cq}. The \texttt{sh:message} captures the CQ in natural language, which helps to identify which question failed to validate from the validation reports.  The \texttt{sh:ask} contains the CQ translated to a SPARQL query prepared in the previous step.

@@

:CQ12 a sh:ConstraintComponent ;
  rdfs:label "Manufacturer-declared test sensitivity"@en ;
  sh:parameter [
    sh:path :cq12
  ] ;
  sh:nodeValidator [
    a sh:SPARQLAskValidator ;
    sh:message """What is the sensitivity of given tests
                  according to their manufacturers?"""@en ;
    sh:prefixes :prefixes ;
    sh:ask """
    ASK {
      $test a act:AntigenCovidTest ;
        schema:manufacturer ?manufacturer .
      [] a ncit:C41394 ;
        dcterms:subject $test ;
        dcterms:creator ?manufacturer ;
        rdf:value ?sensitivity .
    }
    """
  ] .
\end{lstlisting}

\textbf{Step 6:} We started with implementing a transformation of the source data to RDF. We converted the input XML data into RDF/XML via an XSL transformation followed by SPARQL Update operations for post-processing.

\textbf{Step 7:} The resulting data was tested by the CQs implemented in SHACL. We automated the data processing and test execution by a script based on Jena command-line tools.\footnote{\url{https://jena.apache.org/documentation/tools}} Given that this is a small knowledge graph of around 10 thousand RDF triples, we validated it as a whole.

\textbf{Step 8:} The final step of the proposed method is about refactoring artifacts. Our development work proceeded in several iterations guided by continuous feedback from the script for data transformation and validation. When the knowledge graph satisfied the CQs, we continued with refactoring the semantic artifacts we created. For example, we wanted to avoid blank nodes to make the data transformation process deterministic, so we improved the XSL transformation to generate only IRIs instead. We also abstracted a parent class for evaluations in the ontology and adjusted the SHACL shapes accordingly.

Some errors were not detected by the tests implementing the CQs. For example, when querying the knowledge graph, we found that the data transformation to RDF mistakenly created IRIs of manufacturers based on their antigen covid test identifiers. Consequently, the resulting data related each manufacturer to one antigen covid test, so that it was not possible to group multiple tests by the same manufacturer. A common response to finding errors not covered by tests is improving test coverage to enable detecting the errors found and avoid future regressions. We followed this practice and added a CQ comparing a given manufacturer's antigen covid tests. The CQ presupposed that there are manufacturers offering more than one test and would therefore fail in case of the above-described error. Apart from fixing the way we generated manufacturers' IRIs, we spent further effort on de-duplicating manufacturers via post-processing by SPARQL Update operations.

@@

\section{Conclusions}
The hereby proposed method guides through the open-ended process of knowledge graph construction. It breaks the process down into a well-defined sequence of steps, feedback loops between them, and semantic artifacts that are produced or consumed in the process. The central contribution of the method is allowing to test if the produced knowledge graph satisfies the requirements for its construction. This is done via competency questions formulated as SPARQL queries embedded in SHACL shapes for test automation. As such, one of its key advantages is being based on established semantic web standards.

We share the experience of applying the method to build a knowledge graph about antigen covid tests. The method aided in collaborative development of this knowledge graph, allowing its incremental refinement without regressions. Since constructing knowledge graphs from the same user requirements using multiple methods, while comparing their outcomes, is prohibitively expensive, we expect future improvements of the proposed method to come from its applications on knowledge graphs with different requirements.

>>>>

%%%Turtle
\lstdefinelanguage{turtle}
{
  identifierstyle=\color{afterPrefix},    %afterPrefix
  morecomment=[n][\color{iri}]{<}{>}, % iris
  morestring=[b][\color{black}]{\"},  % strings
  % -------------------------- prefixes
  classoffset=1,
  keywordstyle=\color{prefix},
  morekeywords={dcterms, owl, rdfs, rdf, schema, reg, skos, xsd, sh},
  % -------------------------- rdfType
  classoffset=2,
  keywordstyle=\color{purple},
  morekeywords={a},
  % -------------------------- language
  classoffset=3,
  keywordstyle=\color{orange},
  morekeywords={@en}
}

@@

    \item Wrap the queries as SPARQL-based SHACL constraints to automate their execution. SHACL defines a target\footnote{\urlshacltargets} of each constraint. In our context, a target sets the scope of a CQ. The scope of the data graph validated by a given constraint may either encompass the entire knowledge graph or cover its subset. For instance, existentially quantified queries checking the complete data graph can target it by using \texttt{sh:targetNode []}. It is also possible to rewrite a part of a SPARQL query translating a CQ as target selection in SHACL. Alternatively, the prerequisite part of the query can be represented as a SPARQL-based target \cite{SHACLAF2017}. An~entity-centric partitioning of a given knowledge graph can be implemented by extracting concise bounded descriptions~\cite{Stickler2005} of the targeted entities. Some data sources of knowledge graphs may natively partition data to independent subsets, such as API responses or messages from a queue, that can be effectively validated by specific subsets of SHACL constraints. SHACL can also define the passing criteria of SPARQL-based constraints. An \texttt{sh:ask} query is expected to return the boolean \texttt{true}, while any results produced by an \texttt{sh:select} query are treated as violations. Specific expected results can be either hard-coded into the queries or specified via \texttt{sh:hasValue} in case a single value is expected. Note however, that when including the expected results in SPARQL-based constraints, we may run into the limitations of SHACL due to pre-binding of variables.\footnote{\urlshaclprebinding} For example, the \texttt{VALUES} clause, which could represent the expected results of \texttt{SELECT} queries, is not allowed.

@@

    \item Validate the knowledge graph under development with the SHACL constraints. If the validation fails, resolve the reported violations by fixing the artifacts created in the previous steps. Fix the most primary artifact causing the violations, i.e. their root cause. For instance, data transformation shall not work around an insufficiently expressive ontology. Similarly, imposing a more expressive data model on the knowledge graph might reveal errors in its source data. In this way, the finer structure of its ontology makes the previously hidden data quality issues visible. Some feedback may even indicate ill-formed CQs in need of revision.
    \item Refactor the artifacts. The SHACL specification of the knowledge graph is a key artifact to refactor. It can serve as a formal specification of the anticipated use of the ontology. Moreover, SHACL can transcend a single ontology and specify how to combine terms reused from multiple ontologies. During the refactoring, SHACL core constraints can be induced from SPARQL-based constraints to represent the frequently queried patterns in data. For instance, aided by the knowledge about the domain, you can infer universally quantified axioms represented as SHACL constraints from a sample of existentially quantified axioms represented as SPARQL queries. In particular, SHACL can prescribe the expected relations in data. More complex or distant relations might even require the expressivity of SHACL to be represented, using property paths or SPARQL graph patterns. For example, even though SHACL does not support conditional constraints directly, they can be rewritten according to the inference rule of material implication ${P\to Q\Leftrightarrow \neg P\lor Q}$.

@@

\section{Non-functional Requirements}

Note that the above-described method tests if a knowledge graph meets the given user requirements. It does not evaluate how well these requirements are satisfied. Absence of errors does not imply that the knowledge graph is sound. Therefore, the functional user requirements should be complemented by non-functional requirements describing the desired qualities of the solution. These requirements can be checked and refactored by using alternative sources of feedback, such as:

\begin{description}
    \item[Code review] \hfill \\
    Elicit expert insight from ontology and data engineers.
    \item[User feedback] \hfill \\
    The results produced by SPARQL queries formalizing the CQs can be judged to be incorrect by users. User feedback can identify wrong interpretation of CQs or false assumptions. It might prompt revisiting any of the previously described steps or lead to adding more CQs.
    \item[Usability] \hfill \\
    Usability may manifest in developer experience of writing SPARQL queries on a knowledge graph. In particular, complexity and verbosity of SPARQL queries may be considered. In this way, usability of the queries indirectly reflects the usability of the ontology. For example, verbosity may be caused by duplicate data that can be abstracted to the ontology, while complexity can be reduced by introducing ontological shortcuts. Verbose SPARQL queries may also suggest a need for introducing abstraction or additional axioms into the ontology. In fact, some CQs can be answered directly using the knowledge formalized in the ontology, since SHACL expects the validated data graph\footnote{\urlschaclgraphdata} to include the ontologies it populates.
    \item[Performance] \hfill \\
    Performance can be evaluated indirectly using the execution time of SHACL validation. This time is proportional to the execution times of the SPARQL queries the constraints include.
\end{description}

The non-functional requirements can also be covered by tests, such as query performance tests. These tests can provide additional feedback informing the test-driven development. Using the above-mentioned feedback is a balancing act of multi-objective optimization. For example, performance feedback may be incompatible with the feedback about verbosity or CQs may pose mutually conflicting requirements. Addressing the feedback thus requires making informed trade-offs.

\section{Limitations}

We are aware of several limitations of the proposed method. Here we recount some of them and suggest how to remedy them. The method may cause the developed artifacts outlined in the Figure \ref{fig:process} to overfit the user requirements in scope, which would hinder the reuse of the artifacts. This shortcoming can be remedied by including more CQs or focusing on reusability and abstraction during refactoring. User requirements for the knowledge graph under construction can be more complex than what SPARQL can express. One way around it is formalizing the requirements in a more expressive programming language that extends SHACL, such as in \cite{SHACLJS2017}. Ultimately, in order to ameliorate the limitations of this method, it is best combined with other methods, such as those for ontology design (e.g., Kendall and McGuinness \cite{Kendall2019}).

@@

The data was collected manually from several regulatory bodies, such as SÚKL\footnote{\url{https://www.sukl.cz/prehled-testu-k-diagnostice-onemocneni-covid-19}} and EU HSC\footnote{\url{https://health.ec.europa.eu/health-security-and-infectious-diseases/crisis-management/covid-19-diagnostic-tests_en}}, and was combined with performance evaluations of the tests, such as their sensitivity and specificity, that came from independent studies from various countries.
The data was collected into an XML file.
The final destination of the data was originally twofold.
One was the actual portal that provided a graphical visualization of various properties of the tests.\footnote{\url{https://covidtesty.vse.cz/english/test-evaluation-older-data/}}
The other was a dashboard allowing to perform analyses online.
While the XML data was already structured, it lacked any ontological grounding, and some of its features were tuned towards presentation aspects within the website.

@@

\textbf{Step 1:} The first step of the proposed method is gathering user requirements. We started with capturing user requirements formulated as CQs, such as \textit{"What is the sensitivity of given tests according to their manufacturers?"} We came up with 19 CQs in total, covering both basic information look-up and more complex analytical questions. We removed one of the CQs later upon finding that it was subsumed by another CQ. The complete list of the CQs is available online.\footnote{\url{https://github.com/KIZI/antigen-covid-tests-knowledge-graph/wiki/Competency-questions}}

@@

\textbf{Step 4:} Having a minimum viable ontology, we were able to translate the CQs into SPARQL queries, such as in the example Listing \ref{lst:example-cq-sparql}. Each question was translated manually.

\begin{lstlisting}[caption=Example competency question in SPARQL, label={lst:example-cq-sparql}]
PREFIX act:     <https://covidtesty.vse.cz/vocabulary#>
PREFIX dcterms: <http://purl.org/dc/terms/>
PREFIX ncit:    <http://purl.obolibrary.org/obo/NCIT_>
PREFIX rdf:     <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
PREFIX schema:  <http://schema.org/>

ASK {
  ?test a act:AntigenCovidTest ;
    schema:manufacturer ?manufacturer .

  [] a ncit:C41394 ; # Diagnostic sensitivity
    dcterms:subject ?test ;
    dcterms:creator ?manufacturer ;
    rdf:value ?sensitivity .
}
\end{lstlisting}

\textbf{Step 5:} We wrapped the SPARQL queries as SHACL constraints, such as in the example Listing \ref{lst:example-cq} that defines a SPARQL-based constraint component\footnote{\url{https://www.w3.org/TR/shacl/#sparql-constraint-components}} applied to the entire data graph. The \texttt{sh:message} captures the CQ in natural language, which helps to identify which question failed to validate in the validation reports. The \texttt{sh:ask} contains the CQ translated to a SPARQL query prepared in the previous step.

@@

:CQ12 a sh:ConstraintComponent ;
  rdfs:label "Manufacturer-declared test sensitivity"@en ;
  sh:parameter [
    sh:path :cq12
  ] ;
  sh:nodeValidator [
    a sh:SPARQLAskValidator ;
    sh:message """What is the sensitivity of given tests
                  according to their manufacturers?"""@en ;
    sh:prefixes :prefixes ;
    sh:ask """
    ASK {
      ?test a act:AntigenCovidTest ;
        schema:manufacturer ?manufacturer .

      [] a ncit:C41394 ; # Diagnostic sensitivity
        dcterms:subject ?test ;
        dcterms:creator ?manufacturer ;
        rdf:value ?sensitivity .
    }
    """
  ] .
\end{lstlisting}

\textbf{Step 6:} We started with implementing a transformation of the source data to RDF. We converted the input XML data into RDF/XML via an XSL transformation followed by SPARQL Update operations for post-processing. For example, we used post-processing for normalization of code-list values and manufacturers' names.

\textbf{Step 7:} The resulting data was tested by the CQs implemented in SHACL. We automated the data processing and test execution by a shell script based on Jena command-line tools.\footnote{\url{https://jena.apache.org/documentation/tools}} Given that this is a small knowledge graph of around 10 thousand RDF triples, we validated it as a whole.

\textbf{Step 8:} The final step of the proposed method is about refactoring the developed artifacts. Our development work proceeded in several iterations guided by continuous feedback from the script for data transformation and validation. When the knowledge graph satisfied the CQs, we continued with refactoring the semantic artifacts we created. For example, we wanted to avoid blank nodes to make the data transformation results better comparable, so we improved the XSL transformation to generate IRIs instead. We also abstracted a parent class for evaluations in the ontology and adjusted the SHACL shapes accordingly.

Some errors were not detected by the tests implementing the CQs. For example, when querying the knowledge graph, we found that the data transformation to RDF mistakenly created IRIs of manufacturers based on their antigen covid test identifiers. Consequently, the resulting data related each manufacturer to one antigen covid test, so that it was not possible to group multiple tests by the same manufacturer. A common response to discovering errors not covered by tests is to improve the test coverage to enable detecting the errors found and avoid future regressions. We followed this practice and added a CQ comparing a given manufacturer's antigen covid tests. The CQ presupposed that there are manufacturers offering more than one test and would therefore fail in case of the above-described error. Apart from fixing the way we generated manufacturers' IRIs, we spent further effort on de-duplicating manufacturers via post-processing by SPARQL Update operations.

@@

\section{Conclusions}

The hereby proposed method guides through the open-ended process of knowledge graph construction. It breaks the process down into a well-defined sequence of steps, feedback loops between them, and semantic artifacts that are produced or consumed in the process. The central contribution of the method is allowing to test if the produced knowledge graph satisfies the requirements for its construction. This is done via competency questions formulated as SPARQL queries embedded in SHACL shapes for test automation. As such, one of its key advantages is being based on the established semantic web standards.

We shared the experience of applying the method to build a knowledge graph about antigen covid tests. The method aided in collaborative development of this knowledge graph, allowing its incremental refinement without regressions. Since constructing knowledge graphs from the same user requirements using multiple methods, while comparing their outcomes, is prohibitively expensive, we expect future improvements of the proposed method to come from its applications on knowledge graphs with different requirements.
